---
title: "BIOLM0039 - Programming in R coursework"
author: "Archie Benn (1935263)"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Set workspace

```{r, echo = TRUE, message=FALSE, error=FALSE}
# reset environment
rm(list=ls())

# load libraries
library(tidyverse)
library(performance)
library(Hmisc)
library(ggpubr)
library(rstatix)
library(fitdistrplus)
library(MASS)
library(see)
library(TMB)
library(glmmTMB)
library(DHARMa)
library(data.table)
library(caret)
library(modelbased)
library(broom.mixed)
library(dotwhisker)

```

## 2. Import and check Data
Load in the data (tsv file saved from GitHub) as a tibble:

```{r}
df <- read_tsv("../data/mass_sj19031.tsv")

# check class (= data.frame/tibble)
class(df)
```

Describe the data to get an overview of variables, values, and any missing data or typos:

```{r}
describe(df)
```

With this I can see the following:  
- Data collected on giraffe mass from 5 observers across 4 parks  
- BOOL type categorical data for `lion_presence`, `vegetation`, and `landscape`    
- Continuous data on average annual rainfall, distance to water, and predator density  
- Missing values in `landscape`, `vegetation`, `lion_presence`, `env_value`  
- 5 distinct values for the BOOL variable `lion_presence` -> Typos in 'Absent' or 'Present' 

## 2. Tidying data
### 2.1 Outliers
To check outliers and see a general distribution of body mass over the full dataset `df`, histograms and boxplots will be created:

```{r}
# histogram of entire data mass to view distribution
mass_hist <- df %>%
  
  ggplot(aes(x = body_mass)) +
  
  # add histogram with bin width at 250kg
  geom_histogram(fill = "#9EB9F3", color = "black", binwidth = 250) +
  
  # aesthetics
  theme_minimal() +
  labs(title = "Histogram of giraffe body mass across original dataset") +
  xlab("Body mass (kg)") +
  ylab("Count")

mass_hist
```
```{r}
# boxplot of entire dataset mass to check potential outliers
mass_box <- df %>%
  
  ggplot(aes(x = "", y = body_mass)) +
  
  # reduce width to 0.25
  geom_boxplot(width = 0.25, fill = "coral1") +
  
  # aesthetics
  theme_lucid() +
  labs(title = "Box plot of giraffe mass across original dataset") +
  
  # centre title
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Body mass (kg)")

mass_box
```

This box plot identifies 5 potential mass outliers in `df` as shown by the points outside the box plot using the IQR criterion.    

From this information an outlier data frame will be created using `identify_outliers()` from `rstatix` which includes all rows related to these outliers in body mass, and includes outliers and extreme values.   

This data frame will be used to remove these data during cleaning, and is justified biologically as giraffes rarely exceed 2000kg and outliers will be identified statistically:

```{r}
# create outlier data frame
outlier_df <- identify_outliers(data = df,
                  variable = "body_mass")

head(outlier_df)
```

The dataset `df` is also in long format, and to have each row as an individual observation the data will be transformed to wide format, and cleaned again to drop any other N/A values which arise from the transformation to wide data below.

### 2.2 Create cleaned data frame
A clean data frame will be created based on the information above:

```{r, message=FALSE, error=FALSE}
# using tidyverse
df_clean <- df %>%
  
  # remove outliers using anti_join() and the outlier data frame:
  anti_join(outlier_df)  %>%
  
  # transform from long to wide format:
  pivot_wider(names_from = env_var, values_from = env_value) %>%

  # remove NA values identified previously, including those in the new 'wide columns':
  drop_na(landscape, vegetation, lion_presence, distance_to_water_km, average_annual_rainfall_mm, predator_density_per_sq_km) 


# fix typos in lion_presence (no need to remove as very clearly 'Present' or 'Absent' typos)
df_clean$lion_presence[df_clean$lion_presence == "Preesnt"] <- "Present" 
df_clean$lion_presence[df_clean$lion_presence == "Preseet"] <- "Present"
df_clean$lion_presence[df_clean$lion_presence == "Asbent"] <- "Absent"

# check cleaned data 
describe(df_clean)
```

This `clean_df` now has no missing variables, fixed `lion_presence' data, no statistical outiers based on the original `df`, and each row is an individual observation.

## 3. Definition of hypotheses variables
### 3.1 Hypotheses
ADD HERE

### 3.2 Variables
`df_clean` contains variables as follows:  

**Response:**  
- `body_mass`  

**Predictors (fixed):**  
 - Categorical = `landscape`, `vegetation`, `lion_presence`  
 - Continuous = `distance_to_water_km`, `average_annual_rainfall_mm`, and `predator_density_per_sq_km`  
 
**Predictors (random)**
These two groups only have 4 and 5 levels each which could make variance estimates unstable/ They could also be included as fixed effects but to start with they will be included as random effects to account for clustering among observations.
 - Groups/grouping: `park`, `observer`   
 
These data are spatially structured, and no information on temporal structure is provided. Also need to consider if random effects will be considered as slopes or intercepts, and any interactions when forming models.  

The response (body mass) is continuous and $\geq$ 0, which allows for Gaussian and/or Gamma if a GLMM is to be used.
 
### 3.3 Crossed or nested random effects?
If any observer makes observations in $\gt$ 1 park, the random effects should be crossed, but if observers are contained within only one park and do not make observations elsewhere, they can be treatd as nested. A table with both park and observers can show the number of observations for each observer across each park:

```{r}
table(df_clean$observer, df_clean$park)
```

So each observer makes observations at all parks, therefore the random effects must be crossed.  

## 4. Raw data exploration
### 4.1 Viewing summaries for potential random effect variables:
For a summary of giraffe mass across each of the four parks:

```{r}
# mass across parks
mass_parks <- df_clean %>%
  group_by(park) %>%
  summarise(mean_mass = mean(body_mass))

mass_parks
```

And as a violin plot:

```{r}
# violin for mass across parks
plot_mass_parks <- df_clean %>%
  
  ggplot(mapping = aes(x = park, y = body_mass)) +
  
  # violin plot 
  geom_violin(aes(fill = park)) + 
  
  # hide redundant legend
  guides(fill = "none") +
  
  # add boxplot 
  geom_boxplot(width = 0.1, col = "black")+
  
  # aesthetics
  ggtitle("Violin plot of giraffe mass across the parks") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45)) +
  ylab("Body mass (kg)") + 
  xlab("Park") 

plot_mass_parks
```

Zebra Horizon Park appears to have a slightly raised mean giraffe mass copared to the other three parks.  

For the sorted mean across each combination of the variables `landscape`, `lion_presence`, `vegetation`, and `oberver` to get a general idea if any of these appear to have a strong impact over the others in terms of giraffe mass (any particular groupings at the top, for instance):

```{r, message=FALSE, error=FALSE}
mass_lvl <- df_clean %>%
  group_by(landscape, lion_presence, vegetation, observer) %>%
  
  # create mean mass for summary
  summarise(mean_mass = mean(body_mass)) %>%
  
  # sort by descending mean mass
  arrange(desc(mean_mass))

mass_lvl
```

The top four mean mass values are all observer 1, so potential for bias from this observer, but just an idea. 

### 4.2 Scatter plots of continuous variables across each park
Log distance to water:

```{r}
# Log distance to water vs body mas scatter plot
plot_water <- df_clean %>%
  
  # log the distance to water
  ggplot(aes(x=log(distance_to_water_km+ 1), y=body_mass)) +
  
  # plot data points
  geom_point(col="coral1") +
  
  # add linear regression
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y~x, linetype = "dashed") +
  
  # aesthetics
  ggtitle("Scatter plot of giraffe mass against log(distance to water) across each of the parks") +
  xlab("log(Distance to water (km))") +
  ylab("Body mass (kg)") +
  
  # group by park
  facet_wrap(. ~ park) +
  theme_lucid() 

plot_water
```

The distance to water appears to have a slight positive effect on body mass across 3 out of the 4 parks.

Average rainfall:

```{r}
plot_rain <- df_clean %>%
  
  # average rainfall to body mass
  ggplot(aes(x=average_annual_rainfall_mm, y=body_mass)) +
  
  # add data points
  geom_point(col="#9EB9F3") +
  
  # add linear regression
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y~x, linetype = "dashed") +
  
  # aesthetics
  ggtitle("Scatter plot of giraffe mass against average annual rainfall across each of the parks") +
  xlab("Average annual rainfall (mm)") +
  ylab("Body mass (kg)") +
  
  # group by park
  facet_wrap(. ~ park) +
  theme_lucid() 

plot_rain
```

Average annual rainfall appears to have a slight positive effect on body mass in 3 out of the 4 parks.

Predator density:

```{r}
plot_predator <- df_clean %>%
  
  # log the predator density
  ggplot(aes(x=log(predator_density_per_sq_km), y=body_mass)) +
  
  # add data points
  geom_point(col="#F89C74") +
  
  # add linear regression
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y~x, linetype = "dashed") +
  
  # aesthetics
  ggtitle("Scatter plot of giraffe mass against log(predator density) across each of the parks") +
  xlab("log(Predator density per square kilometre)") +
  ylab("Body mass (kg)") +
  
  # group by park
  facet_wrap(. ~ park) +
  theme_lucid() 

plot_predator
```

Predator density doesn't appear to have a strong relationship with body mass across the parks.


### 4.3 Distribution and normality checks
The response variable is body mass of the giraffes. This is a continuous and strictly positive, and as such the main families involved for a model are likely Gaussian and Gamma. Gaussian requires a normal distribution, while Gamma typically fits skewed data more reliably. To suggest a family QQ plots and histograms will be plotted.  

Histogram with overlaid normal of body mass in clean data:

```{r, message=FALSE, error=FALSE}
mass_hist_clean <- df_clean %>%
  
  ggplot(aes(x = body_mass)) +
  
  # add histogram with bin width at 50kg and scale using ..density..
  geom_histogram(aes(y = ..density..), fill = "coral1", color = "black", binwidth = 50) +
  
  # add a normal distribution to plot 
  stat_function(fun = dnorm,
                args = list(
                  mean = mean(df_clean$body_mass),
                  sd = sd(df_clean$body_mass)
                  )) +
  
  # aesthetics
  theme_minimal() +
  labs(title = "Histogram of giraffe body mass") +
  xlab("Body mass (kg)") +
  ylab("Density") 

mass_hist_clean
```

Histogram indicates a right skewed body mass in this cleaned data, but approximately normal.

Test QQ-plot across body mass in clean dataset:

```{r}
ggqqplot(df_clean$body_mass,
         title = "QQ plot of body mass")
```

This QQ plot is skewed with tailed a end at lower masses, indicating a right skew and less extreme values at lower masses than expected compared to the reference line. Log transormation of the body mass will be attampted next.

```{r}
# create log body mass column
df_clean <- df_clean %>%
  mutate(log_body_mass = log(body_mass))

# check if column has been created
head(df_clean$log_body_mass)
```

Histogram of log transformed body mass:

```{r, message=FALSE, error=FALSE}
mass_hist_clean_log <- df_clean %>%
  
  ggplot(aes(x = log_body_mass)) +
  
  # add histogram with bin width at 50kg and scale using ..density..
  geom_histogram(aes(y = ..density..), fill = "#154360", color = "black", binwidth = 0.05) +
  
  # add a normal distribution to plot 
  stat_function(fun = dnorm,
                args = list(
                  mean = mean(df_clean$log_body_mass),
                  sd = sd(df_clean$log_body_mass)
                  )) +
  
  # aesthetics
  theme_minimal() +
  labs(title = "Histogram of log giraffe body mass") +
  xlab("Log dody mass (kg)") +
  ylab("Density") 

mass_hist_clean_log
```

Using the natural logarithm of the body mass approximates a normal distribution more closely than the raw body mass.  

QQ plot for these transformed data:

```{r}
ggqqplot(df_clean$log_body_mass,
         title = "QQ plot of log-transformed body mass")
```

This appears to be a more normally distributed QQ plot with a more compressed upper tail. A Box-Cox transformation from `MASS` will be used to calculate a log-likelihood number which best approximates normal using these body mass data: 

```{r}
# first creat a linear model of the body mass, intercept only
lm_body_mass  <- lm(body_mass ~ 1, df_clean)

# apply box-cos to generate a log-likelihood plot which tests lambda form -2 to 2, step 0.1
boxcox(lm_body_mass, lambda = seq(-2, 2, 0.1))
```

From this the highest log-likelihood appears to be approximately 0.25 which is ~ a cube root (1/3). Will attempt a cube root transformation too:

```{r}
# create log body mass column
df_clean <- df_clean %>%
  
  # add cube root transformed column
  mutate(cube_body_mass = body_mass^(1/3))

# check if column has been created
head(df_clean$cube_body_mass)
```

QQ plot for cube root transformed mass: 

```{r}
ggqqplot(df_clean$cube_body_mass,
         title = "QQ plot of cube root-transformed body mass")
```

This cube root-transformed data appears to approximate normal slightly better than the raw mass and log-transformed mass. However these changes are small and do not reflect how the residuals will turn out after creating the model, so raw and the above two transformations will be included in model creation. 

### 4.4 Family checks
Will use `fitdistr` to check family fits.

```{r}
# normal with raw body mass
plot(fitdist(df_clean$body_mass, "norm"))

# gamma with raw body mass
plot(fitdist(df_clean$body_mass, "gamma"))

# normal with log-transformed body mass
plot(fitdist(df_clean$log_body_mass, "norm"))

# normal with cube root-transformed body mass
plot(fitdist(df_clean$cube_body_mass, "norm"))
```

All of these plots show that the data (raw and transformed) are able to fit Gaussian and Gamma (for raw data) well. Will move on to fitting candidate models and then testing these.


## 5. Model formation
Generalised Linear Mixed (effect) Models (GLMMs) will be created here using the `glmmTMB` function. Multiple models will be formed before undergoing intital residual and model testing, followed by statistical comparisons between models to select the most appropriate to take forward.

### Log-transformed Gaussian Model
The first model will be a log-transformed response GLMM with the Gaussian family (identity link), including all predictor variables to capture potential effects of predictors on the response. Starting from log transformed response was determined due to right-skewed raw data.

Random effects will be included for park and observer as intercepts and not as random slopes, as the effects of the predictors on the response is assumed to be the same across levels within groups. Also with a small number of levels in each group (4 and 5 only), using random slopes could be unstable while adding complexity to the model. 


```{r, error=FALSE, message=FALSE}
# fit random effects model
glmm_gauss_log <- glmmTMB(

  # log transformed response ~ fixed effects:
  log(body_mass) ~ landscape +
    vegetation +
    lion_presence +
    distance_to_water_km +
    average_annual_rainfall_mm +
    predator_density_per_sq_km + 
    
    # crossed random effects (park and observer), each with its own intercept:
    (1 | park) +
    (1 | observer),
  
  data = df_clean,
  
  family = gaussian()
)

summary(glmm_gauss_log)
r2_nakagawa(glmm_gauss_log)
```

#### Model check 1: Residual structure with `DHARMa`:

```{r}
# simulate residuals for model with 1000 simulations to reduce stochastic error and stabilise values:
glmm_gauss_log_res <- simulateResiduals(glmm_gauss_log, n =1000)
plot(glmm_gauss_log_res)
```

(comment on residuals plot)

#### Model check 2: Multicollinearity between predictors:

```{r}
check_collinearity(glmm_gauss_log)
```


#### Model check 3: Overdispersion checks

```{r}
check_overdispersion(glmm_gauss_log)
```

#### Model check 4: `check_model` overview:

```{r, fig.width=10, fig.height=10}
check_model(glmm_gauss_log)
```

### Fitting two further models

```{r, message=FALSE, error=FALSE, warning=FALSE}
# create gamma model (without log transformed response)
glmm_gamma <- update(glmm_gauss_log, formula = body_mass ~ ., family = Gamma(link = "log"))
glmm_gauss_cubed <- update(glmm_gauss_log, formula = cube_body_mass ~ .)
```

These two models also underwent the same tests as the first model above, and both fit well, with performance comparable to `the first model`glmm_gauss_log`.

## 6. Model testing
### 6.1 Model predictions visualised
A visualistaion of the above tests with plotting of observed mass vs predicted mass from each model, using `modelbased`:

```{r, message=FALSE, error=FALSE}
# get predicted data frame for each model based on the observed predictor values at each observation using estimate_expectation()
pred_gamma <- estimate_expectation(glmm_gamma)
pred_gauss_log <- estimate_expectation(glmm_gauss_log)
pred_gauss_cubed <- estimate_expectation(glmm_gauss_cubed)

# create df of predicted values based off models' body mass (stored as 'Predicted')
models_predictions <- tibble(
  observed = df_clean$body_mass,
  predicted_gamma = pred_gamma$Predicted,
  
  # back transform predictions from log and cubed models:
  predicted_gauss_log = exp(pred_gauss_log$Predicted),
  predicted_gauss_cubed = (pred_gauss_cubed$Predicted)^3
)

# create observed vs. predicted plot for all models:
models_predictions %>%
  
  ggplot() +
  # add points from each model including gauss2 model
  geom_smooth(aes(x = observed, y = predicted_gamma, colour = "gamma"), method = "lm", se = FALSE) +
  geom_smooth(aes(x = observed, y = predicted_gauss_log, colour = "gauss_log"), method = "lm", se = FALSE) +
  geom_smooth(aes(x = observed, y = predicted_gauss_cubed, colour = "gauss_cubed") , method = "lm", se = FALSE) +
  
  # add slope = 1 line for observed = predicted
  geom_abline(aes(intercept = 0, slope = 1, colour = "1:1"),  linetype = "dashed",) +  
  
  # adding a legend 
  scale_colour_manual(name = NULL,
                     breaks = c("gamma", "gauss_log", "gauss_cubed", "1:1"),
                     labels = c("Gamma Model", "Log Gauss Model", "Cube-root Gauss Model", "Observed = Predicted"),
                     values = c("#1ABC9C", "#154360", "#FFC300", "black")) +
  
  ggtitle("Observed vs Predicted body mass for each model") +
  ylab("Body mass (predicted)") +
  xlab("Body mass (observed)") +
  theme_lucid()

```

As the dashed line is a 1:1 observed set of values, clearly all three models are over fitting at lower masses, and under fitting at higher masses, despite passing residual and other model checks earlier.

### 6.2 `ANOVA` model comparisons
As there are currently three working models which have all passed `DHARMa` residual and `check_model()` checks and appear to predict very similar body masses, their fits can be directly quanitified against each other using `anova()` with `AIC` and `BIC` acting as comparative scores. 

AIC and BIC try to balance how well a model fits with its complexity (models with more parameters are penalised harder to avoid overfitting). A lower score indicates a batter model in both AIC and BIC, with BIC penalising complexity more than AIC:  

```{r}
anova(glmm_gauss_log, glmm_gamma, glmm_gauss_cubed)
```

`glmm_gauss_cubed` has very high deviance and a very significant Chi-squared statistic ($p \lt 2e^{-16}$, indicating a very poor fit to the data. In both AIC and BIC `glmm_gauss_log` scored the lowest of all models and as such will be taken forward as the candidate model.

## 7. Model refining
With `glmm_gauss_log` selected as the candidate model, model parameters will next be refined and further testing will take place, along with visualisation, in order to determine the most representative model for these data.

Another summary of `glmm_gauss2`:
```{r}
summary(glmm_gauss_log)
```


### 7.1 Model simplifying
From the summary above only `distance_to_water_km` is a significant predictor in giraffe body mass in this model, and there are two non-significant predictors: `preditor_density_per_sq_km` ($p = 0.956$) and `lion_presencePresent` ($p = 0.452$). Also, the `Estimate` for average rainfall is so small ($\approx$ 5.7e-5) it's  negligible in this model. 

A model will be made without these three predictors to test if reducing complexity still produces a working model:

```{r}
# update candidate model to remove the lion/predator fixed effects:
glmm_gauss2_log <- update(glmm_gauss_log, formula. = . ~ . - average_annual_rainfall_mm - lion_presence - predator_density_per_sq_km)

# check model summary:
 summary(glmm_gauss2_log)
```

Look at tidy summary of effects:

```{r}
broom.mixed::tidy(glmm_gauss2_log, conf.int = TRUE)
```

Compare `glmm_gauss_log` and `glmm_gauss2_log`:

```{r}
anova(glmm_gauss_log, glmm_gauss2_log)
```

Removing the above fixed effects did not significantly worsen the model fit ($p = 0.456$), and the lower AIC/BIC vales mean this model is preferred for parsimony.

Check model residuals with DHARMa, and `check_model()`:  

```{r}
# simulate residuals with 1000 simulations:
glmm_gauss2_log_res <- simulateResiduals(glmm_gauss2_log, n = 1000)
plot(glmm_gauss2_log_res)
```
```{r, fig.width=10, fig.height=10}
check_model(glmm_gauss2_log)
```

Still passes residual testing and other diagnostics. As such the `glmm_gauss2_log` model will remain and be used as the final model.

## 8. Model fit and interpretation

This model predictions can be plotted against the observed values to visualise how well it fits the observed data:

```{r}
# create predicted and observed data frame to get predicted values for mass from the model based on the predictor values at each true observation:
fit_model_gauss2_log <- data.frame(
  
  # back transform from log to get predicted on the correct scale:
  "Predicted" = exp(predict(glmm_gauss2_log, type = "response")),
  "Observed" = df_clean$body_mass
  
)

# plot predicted vs observed with 1:1 (observed:observed) comparison line
ggplot(fit_model_gauss2_log) +
  
  # each point (x) = observed, (y) = model prediction
  geom_point(aes(x = Observed, y = Predicted,)) +
  
  # add linear regression of observed:predicted for model
  geom_smooth(
    aes(x = Observed, y = Predicted, colour = "gauss"),
    method = lm,
    se = FALSE) +
  
  # add slope = 1 line for observed = predicted
  geom_abline(aes(intercept = 0, slope = 1, colour = "1:1"),  linetype = "dashed",) +

  # aesthetics
  scale_colour_manual(name = NULL,
                      breaks = c("1:1", "gauss"),
                      labels = c("Observed = Predicted", "Log Gauss2 Model"),
                      values = c("black", "#FF5733")) +
  
  xlab("Body mass (observed)") +
  ylab("Body mass (predicted)") +
  
  ggtitle("Predicted vs. Observed Giraffe Body Mass for Log Gauss2 Model") +
  theme_lucid()

```

Points represent the predicted mass value from the `glmm_gauss2_log` model based on the predictor values from the true observations in the data. The fitted line shows that the model tends to over predicts lower body masses, and under predict higher body masses compared to the observed values, but overall residuals and model checks indicate a good fit. As the main aim is to test any significant predictors on body mass, this model is appropriate.

Variance in the observed response (log(body mass)) captured by the model vs. residual error can be quantified using `r2_nakagawa()` from `performance`:

```{r}
r2_nakagawa(glmm_gauss2_log)
```

From this, the conditional r-squared (random and fixed effects in the model) only capture 22.1% of the variance, with marginal r-squared (fixed effects only) capturing 2.7% of the variance. Therefore the random effects account for about 19.5% of the variance in the response.  This means that the remaining 77.9% of the variance cannot be explained by the model which could be caused by differences not accounted for by the chosen predictors, or other biological factors not included in the model. 


## 9. Model equation and predictor significance
While the exact coefficients are not strictly necessary for this analysis as it is just significance of any predictors on body mass being tested, however it can be helpful to understand the model in its linear and back transformed form.

### 9.1 Model equation

Summary of `glmm_gauss2_log`:

```{r}
summary(glmm_gauss2_log)
```

Therefore `glmm_gauss2_log` can be expressed as:  
$$
\text{log(Body Mass}_{i}\text{(kg))} = \beta_0 + \beta_1\text{Landscape}_{i} + \beta_2\text{Vegetation}_{i} + \beta_3\text{Distance to Water}_{i} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}
\tag{1}
$$

Where:  

- $i$ represents the observation
- $\beta_0$ is the expected body mass when all other predictors are at reference level or mean values.  
- $\beta$ values are the model variable coefficients  
- $\text{Landscape}_{i}$ -> BOOL landscape value: hilly (=1) or flat (=0)  
- $\text{Vegetation}_{i}$ -> BOOL vegetation value: open (=1) or dense (=0)  
- $\text{Distance to Water}_{i}$ is the distance to water (km) 
- $b_{park[i]}$ is the intercept shift for park in observation $i$  
- $b_{observer[i]}$ is the intercept shift for observer in observation $i$  
- $\epsilon_{i}$ is the error term (for residual error)  

With coefficients: 
$$
\text{log(Body Mass}_{i}\text{(kg))} = 7.09 - 0.041\cdot\text{Landscape}_{i}\text{} - 0.039\cdot\text{Vegetation}_{i} + 0.012\cdot\text{Distance to Water}_{i}\text{(km)} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}
$$

### 9.2 Back transformation 
Back transformation gets from log (natural) to the normal scale:  

$$
\text{Body Mass}_{i}\text{(kg)} = e^{7.09 - 0.041\cdot\text{Landscape}_{i}\text{} - 0.039\cdot\text{Vegetation}_{i} + 0.012\cdot\text{Distance to Water}_{i}\text{(km)} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}}
$$

On this scale, the predictors now have multiplicative, rather than additive, effects on the body mass in this model. In other words any coefficient $\beta$ on the log scale now has an $e^{\beta}$ multiplicative effect on body mass.

### 9.3 Significant predictor: distance from water
This model suggests that the only significant predictor on giraffe body mass is the Distance from water ($p = 0.025$). The effect of this predictor can be written as: 

$$
\text{Body Mass}_i\text{(kg)} \propto e^{0.012\cdot\text{Distance to Water}_{i}\text{(km)}} \approx 1.012^{\text{Distance to Water (km)}}
$$


So the model predicts a small marginal effect: giraffe mass increases by $\approx$ 1.2% for every km further from water the giraffe is when observed, all other predictors constant, and at a significance of $p = 0.025$. 



## 10. Iterpreting model fixed effects

```{r}
# data frame with fixed effect and confidence intervals:
model_fixed_effects_tidy <- broom.mixed::tidy(glmm_gauss2_log, conf.int = TRUE) %>%
  
  # don't include random effects (see next section):
  filter(effect == "fixed") %>%
  
  # remove intercept value
  filter(term!="(Intercept)")
  


# back transform estimates and confidences:
model_fixed_effects_tidy <- model_fixed_effects_tidy %>%
  
  # back transform
  mutate(
    estimate_bt = exp(estimate),
    conf.low_bt = exp(conf.low),
    conf.high_bt = exp(conf.high)
  )


  
 model_fixed_effects_tidy 



ggplot(model_fixed_effects_tidy, aes(x=estimate, y=term)) + 
  
  geom_point() +
  
  geom_errorbar(aes(xmin=conf.low, xmax=conf.high),                       
              width=0.3, 
              col="grey50") +
  
  theme_lucid()


```





## 11. Interpreting model random effects 
From the above equation for `glmm_gauss2_log` (equation 1), there are random effects termed $b_{park[i]}$ and $b_{observer[i]}$. These are estimated multiplicative factors on the body mass (as additive in log-linear) from the model, and they vary between the levels within the random effects.  

To visualise the estimated multiplicative effect for each level, `ranef()` from `glmmTMB` will be used to extract random effects (park and observer):

### 11.1 Park

```{r}
# extract park random effects  with condVar attributes attached (as a list):
ranefs <- ranef(glmm_gauss2_log, condVar = TRUE)

# ranef contains the data frames within $cond, so to extract and mutate park df:
ranefs_park <- ranefs$cond$park %>%
  
  mutate(
    # make each row name a factored column for the data frames
    park = factor((rownames(.))),
    
    # back transform the intercept to get on the normal scale
    bt = exp(`(Intercept)`),
    
    # extract conditional variance from the 3d array it is stored in in this df:
    condVar = attr(., "condVar")[1, 1, ],
    
    # calculate (log) standard errors from condVar:
    se = sqrt(condVar),
  
    # add upper and lower bounds at 95% confidence intervals (1.96 x se)
    bt_lower = exp(`(Intercept)` - 1.96*se),
    bt_upper = exp(`(Intercept)` + 1.96*se)
  )



head(ranefs_park)
```

Now the data frame has the necessary estimated multiplicative factors (`bt`) for each level, they can be plotted along with the errors for each estimate. Note these factors are equivalent to $e^{b_{park[i]}}$ due to back transformation.

```{r, message=FALSE, error=FALSE}
# plot park random effect impact on body mass 

ggplot(ranefs_park, aes(x = bt, y = park)) +
  
  geom_point(col = "#F89C74", size = 5) + 
  
  # add effect horizontal error bars
  geom_errorbarh(aes(xmin = bt_lower, xmax = bt_upper), height = 0.25, col = "grey50") +
  
  # add vertical line for mean/expected across all parks
  geom_vline(xintercept = 1, col ="red3", linetype = "dashed") +
  
  ylab("Park") +
  xlab("Multiplicative factor on body mass ") +
  ggtitle("Estimated Impact of Park (Random Effect) on Body Mass") +
  
  theme_lucid()
```

Error bars here represent a 95% confidence interval after back transforming, and the red dashed line represents the expected body mass across the population after accounting for the fixed and other random effects (as 1 represents no effect on the multiplicative scale for body mass).

Park multiplicative factors, at 95% confidence, range from ~0.88 to ~1.20, and indicate that some parks have a higher/lower baseline body mass. The 95% confidence intervals of Zebra Horizon Park do no cross the mean line, indicating that masses above the population average are reported within this park in the model.

To assess whether park random effects contribute significantly to this model, a model without the parks will be fitted and then compared to `glmm_gauss2_log` with anova:

```{r}
# new model without park random effect:
glmm_no_park <- update(glmm_gauss2_log, ~ . - (1 | park))
# compare with anova:
anova(glmm_gauss2_log, glmm_no_park)
```

The above shows higher AIC/BIC values without the park, and a very significant chi squared value ($p = 2.21e-5$). This indicates meaningful body mass variation across parks and its inclusion improves model fit.


#### 11.2 Observer
And for the observer random effect, the same can be done:  

```{r}
# and for observer (same as park above):
ranefs_observer <- ranefs$cond$observer %>%
  mutate(
    observer = factor((rownames(.))),
    bt = exp(`(Intercept)`),
    condVar = attr(., "condVar")[1, 1, ],
    se = sqrt(condVar),
    bt_lower = exp(`(Intercept)` - 1.96*se),
    bt_upper = exp(`(Intercept)` + 1.96*se)
  )



head(ranefs_observer)
```


```{r, message=FALSE, error=FALSE}
# plot observer random effect impact on body mass 
ggplot(ranefs_observer, aes(x = bt, y = observer)) +
  
  geom_point(col = "#154360", size = 5) + 
  
  # add effect horizontal error bars
  geom_errorbarh(aes(xmin = bt_lower, xmax = bt_upper), height = 0.25, col = "grey50") +
  
  # add vertical line for mean/expected across all observers
  geom_vline(xintercept = 1, col ="red3", linetype = "dashed") +
  
  ylab("Observer") +
  xlab("Multiplicative factor on body mass") +
  ggtitle("Estimated Impact of Observer (Random Effect) on Body Mass") +
  
  theme_lucid()
```

Error bars here represent a 95% confidence interval after back transforming, and the red dashed line represents the expected body mass across the population after accounting for the fixed and other random effects (as 1 represents no effect on the multiplicative scale for body mass).

Observer multiplicative factors, at 95% confidence, range from ~0.83 to ~1.20, indicating a similar estimated body mass effect level to that see across parks, but across observers, and also visually demonstrates an estimated difference in baseline body mass across the observers. The 95% confidence intervals of observer 1 do no cross the mean line, indicating that observer 1 tends to record masses above the population average in this model.

To assess whether observer random effects contribute significantly to this model, a model without the observers will be fitted, as above with parks:

```{r}
# new model without observer random effect:
glmm_no_observer <- update(glmm_gauss2_log, ~ . - (1 | observer))
# compare with anova:
anova(glmm_gauss2_log, glmm_no_observer)
```

As with parks, including observer as a random effect improves model fit significantly ($p = 2.88e-6$), indicating meaningful body mass variation across observers within the model. 

## 12 Conclusions
Distance to water

