---
title: "BIOLM0039 - Programming in R coursework"
author: "Archie Benn (1935263)"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Set workspace

```{r, echo = TRUE, message=FALSE, error=FALSE}
# reset environment
rm(list=ls())

# load libraries
library(tidyverse)
library(performance)
library(Hmisc)
library(ggpubr)
library(rstatix)
library(fitdistrplus)
library(MASS)
library(see)
library(TMB)
library(glmmTMB)
library(DHARMa)
library(data.table)
library(caret)
library(modelbased)
```

## 2. Import and check Data
Load in the data (tsv file saved from GitHub) as a tibble:

```{r}
getwd()
```

```{r}
df <- read_tsv("../data/mass_sj19031.tsv")

# check class (= data.frame/tibble)
class(df)
```

Describe the data to get an overview of variables, values, and any missing data or typos:

```{r}
describe(df)
```

With this I can see the following:  
- Data collected on giraffe mass from 5 observers across 4 parks  
- BOOL type categorical data for `lion_presence`, `vegetation`, and `landscape`    
- Continuous data on average annual rainfall, distance to water, and predator density  
- Missing values in `landscape`, `vegetation`, `lion_presence`, `env_value`  
- 5 distinct values for the BOOL variable `lion_presence` -> Typos in 'Absent' or 'Present' 

## 2. Tidying data
### 2.1 Outliers
To check outliers and see a general distribution of body mass over the full dataset `df`, histograms and boxplots will be created:

```{r}
# histogram of entire data mass to view distribution
mass_hist <- df %>%
  
  ggplot(aes(x = body_mass)) +
  
  # add histogram with bin width at 250kg
  geom_histogram(fill = "#9EB9F3", color = "black", binwidth = 250) +
  
  # aesthetics
  theme_minimal() +
  labs(title = "Histogram of giraffe body mass across original dataset") +
  xlab("Body mass (kg)") +
  ylab("Count")

mass_hist
```
```{r}
# boxplot of entire dataset mass to check potential outliers
mass_box <- df %>%
  
  ggplot(aes(x = "", y = body_mass)) +
  
  # reduce width to 0.25
  geom_boxplot(width = 0.25, fill = "coral1") +
  
  # aesthetics
  theme_lucid() +
  labs(title = "Box plot of giraffe mass across original dataset") +
  
  # centre title
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Body mass (kg)")

mass_box
```

This box plot identifies 5 potential mass outliers in `df` as shown by the points outside the box plot using the IQR criterion.    

From this information an outlier data frame will be created using `identify_outliers()` from `rstatix` which includes all rows related to these outliers in body mass, and includes outliers and extreme values.   

This data frame will be used to remove these data during cleaning, and is justified biologically as giraffes rarely exceed 2000kg and outliers will be identified statistically:

```{r}
# create outlier data frame
outlier_df <- identify_outliers(data = df,
                  variable = "body_mass")

head(outlier_df)
```

The dataset `df` is also in long format, and to have each row as an individual observation the data will be transformed to wide format, and cleaned again to drop any other N/A values which arise from the transformation to wide data below.

### 2.2 Create cleaned data frame
A clean data frame will be created based on the information above:

```{r, message=FALSE, error=FALSE}
# using tidyverse
df_clean <- df %>%
  
  # remove outliers using anti_join() and the outlier data frame:
  anti_join(outlier_df)  %>%
  
  # transform from long to wide format:
  pivot_wider(names_from = env_var, values_from = env_value) %>%

  # remove NA values identified previously, including those in the new 'wide columns':
  drop_na(landscape, vegetation, lion_presence, distance_to_water_km, average_annual_rainfall_mm, predator_density_per_sq_km) 


# fix typos in lion_presence (no need to remove as very clearly 'Present' or 'Absent' typos)
df_clean$lion_presence[df_clean$lion_presence == "Preesnt"] <- "Present" 
df_clean$lion_presence[df_clean$lion_presence == "Preseet"] <- "Present"
df_clean$lion_presence[df_clean$lion_presence == "Asbent"] <- "Absent"

# check cleaned data 
describe(df_clean)
```

This `clean_df` now has no missing variables, fixed `lion_presence' data, no statistical outiers based on the original `df`, and each row is an individual observation.

## 3. Definition of variables
`df_clean` contains variables as follows:  

**Response:**  
- giraffe body mass  

**Predictors (fixed):**  
 - Categorical = landscape, vegetation, lion presence  
 - Continuous = distance to water, rainfall, and predator density  
 
**Predictors (random)**
 - Groups/grouping: park, observer   
 
These data are spatially structured, and no information on temporal structure is provided. Also need to consider which random effects will be considered as slopes or intercepts, and any interactions when forming models.  

The response (body mass) is continuous and $\geq$ 0, which allows for Gaussian and/or Gamma if a GLMM is to be used.
 

## 4. Raw data exploration
### 4.1 Viewing summaries for potential random effect variables:
For a summary of giraffe mass across each of the four parks:

```{r}
# mass across parks
mass_parks <- df_clean %>%
  group_by(park) %>%
  summarise(mean_mass = mean(body_mass))

mass_parks
```

And as a violin plot:

```{r}
# violin for mass across parks
plot_mass_parks <- df_clean %>%
  
  ggplot(mapping = aes(x = park, y = body_mass)) +
  
  # violin plot 
  geom_violin(aes(fill = park)) + 
  
  # hide redundant legend
  guides(fill = "none") +
  
  # add boxplot 
  geom_boxplot(width = 0.1, col = "black")+
  
  # aesthetics
  ggtitle("Violin plot of giraffe mass across the parks") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45)) +
  ylab("Body mass (kg)") + 
  xlab("Park") 

plot_mass_parks
```

Zebra Horizon Park appears to have a slightly raised mean giraffe mass copared to the other three parks.  

For the sorted mean across each combination of the variables `landscape`, `lion_presence`, `vegetation`, and `oberver` to get a general idea if any of these appear to have a strong impact over the others in terms of giraffe mass (any particular groupings at the top, for instance):

```{r, message=FALSE, error=FALSE}
mass_lvl <- df_clean %>%
  group_by(landscape, lion_presence, vegetation, observer) %>%
  
  # create mean mass for summary
  summarise(mean_mass = mean(body_mass)) %>%
  
  # sort by descending mean mass
  arrange(desc(mean_mass))

mass_lvl
```

The top four mean mass values are all observer 1, so potential for bias from this observer, but just an idea. 

### 4.2 Scatter plots of continuous variables across each park
Log distance to water:

```{r}
# Log distance to water vs body mas scatter plot
plot_water <- df_clean %>%
  
  # log the distance to water
  ggplot(aes(x=log(distance_to_water_km+ 1), y=body_mass)) +
  
  # plot data points
  geom_point(col="coral1") +
  
  # add linear regression
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y~x, linetype = "dashed") +
  
  # aesthetics
  ggtitle("Scatter plot of giraffe mass against log(distance to water) across each of the parks") +
  xlab("log(Distance to water (km))") +
  ylab("Body mass (kg)") +
  
  # group by park
  facet_wrap(. ~ park) +
  theme_lucid() 

plot_water
```

The distance to water appears to have a slight positive effect on body mass across 3 out of the 4 parks.

Average rainfall:

```{r}
plot_rain <- df_clean %>%
  
  # average rainfall to body mass
  ggplot(aes(x=average_annual_rainfall_mm, y=body_mass)) +
  
  # add data points
  geom_point(col="#9EB9F3") +
  
  # add linear regression
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y~x, linetype = "dashed") +
  
  # aesthetics
  ggtitle("Scatter plot of giraffe mass against average annual rainfall across each of the parks") +
  xlab("Average annual rainfall (mm)") +
  ylab("Body mass (kg)") +
  
  # group by park
  facet_wrap(. ~ park) +
  theme_lucid() 

plot_rain
```

Average annual rainfall appears to have a slight positive effect on body mass in 3 out of the 4 parks.

Predator density:

```{r}
plot_predator <- df_clean %>%
  
  # log the predator density
  ggplot(aes(x=log(predator_density_per_sq_km), y=body_mass)) +
  
  # add data points
  geom_point(col="#F89C74") +
  
  # add linear regression
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y~x, linetype = "dashed") +
  
  # aesthetics
  ggtitle("Scatter plot of giraffe mass against log(predator density) across each of the parks") +
  xlab("log(Predator density per square kilometre)") +
  ylab("Body mass (kg)") +
  
  # group by park
  facet_wrap(. ~ park) +
  theme_lucid() 

plot_predator
```

Predator density doesn't appear to have a strong relationship with body mass across the parks.


### 4.3 Distribution and normality checks
The response variable is body mass of the giraffes. This is a continuous and strictly positive, and as such the main families involved for a model are likely Gaussian and Gamma. Gaussian requires a normal distribution, while Gamma typically fits skewed data more reliably. To suggest a family QQ plots and histograms will be plotted.  

Histogram with overlaid normal of body mass in clean data:

```{r, message=FALSE, error=FALSE}
mass_hist_clean <- df_clean %>%
  
  ggplot(aes(x = body_mass)) +
  
  # add histogram with bin width at 50kg and scale using ..density..
  geom_histogram(aes(y = ..density..), fill = "coral1", color = "black", binwidth = 50) +
  
  # add a normal distribution to plot 
  stat_function(fun = dnorm,
                args = list(
                  mean = mean(df_clean$body_mass),
                  sd = sd(df_clean$body_mass)
                  )) +
  
  # aesthetics
  theme_minimal() +
  labs(title = "Histogram of giraffe body mass") +
  xlab("Body mass (kg)") +
  ylab("Density") 

mass_hist_clean
```

Histogram indicates a right skewed body mass in this cleaned data, but approximately normal.

Test QQ-plot across body mass in clean dataset:

```{r}
ggqqplot(df_clean$body_mass,
         title = "QQ plot of body mass")
```

This QQ plot is skewed with tailed a end at lower masses, indicating a right skew and less extreme values at lower masses than expected compared to the reference line. Log transormation of the body mass will be attampted next.

```{r}
# create log body mass column
df_clean <- df_clean %>%
  mutate(log_body_mass = log(body_mass))

# check if column has been created
head(df_clean$log_body_mass)
```

Histogram of log transformed body mass:

```{r, message=FALSE, error=FALSE}
mass_hist_clean_log <- df_clean %>%
  
  ggplot(aes(x = log_body_mass)) +
  
  # add histogram with bin width at 50kg and scale using ..density..
  geom_histogram(aes(y = ..density..), fill = "#154360", color = "black", binwidth = 0.05) +
  
  # add a normal distribution to plot 
  stat_function(fun = dnorm,
                args = list(
                  mean = mean(df_clean$log_body_mass),
                  sd = sd(df_clean$log_body_mass)
                  )) +
  
  # aesthetics
  theme_minimal() +
  labs(title = "Histogram of log giraffe body mass") +
  xlab("Log dody mass (kg)") +
  ylab("Density") 

mass_hist_clean_log
```

Using the natural logarithm of the body mass approximates a normal distribution more closely than the raw body mass.  

QQ plot for these transformed data:

```{r}
ggqqplot(df_clean$log_body_mass,
         title = "QQ plot of log-transformed body mass")
```

This appears to be a more normally distributed QQ plot with a more compressed upper tail. A Box-Cox transformation from `MASS` will be used to calculate a log-likelihood number which best approximates normal using these body mass data: 

```{r}
# first creat a linear model of the body mass, intercept only
lm_body_mass  <- lm(body_mass ~ 1, df_clean)

# apply box-cos to generate a log-likelihood plot which tests lambda form -2 to 2, step 0.1
boxcox(lm_body_mass, lambda = seq(-2, 2, 0.1))
```

From this the highest log-likelihood appears to be approximately 0.25 which is ~ a cube root (1/3). Will attempt a cube root transformation too:

```{r}
# create log body mass column
df_clean <- df_clean %>%
  
  # add cube root transformed column
  mutate(cube_body_mass = body_mass^(1/3))

# check if column has been created
head(df_clean$cube_body_mass)
```

QQ plot for cube root transformed mass: 

```{r}
ggqqplot(df_clean$cube_body_mass,
         title = "QQ plot of cube root-transformed body mass")
```

This cube root-transformed data appears to approximate normal slightly better than the raw mass and log-transformed mass. However these changes are small and do not reflect how the residuals will turn out after creating the model, so raw and the above two transformations will be included in model creation. 

### 4.4 Family checks
Will use `fitdistr` to check family fits.

```{r}
# normal with raw body mass
plot(fitdist(df_clean$body_mass, "norm"))

# gamma with raw body mass
plot(fitdist(df_clean$body_mass, "gamma"))

# normal with log-transformed body mass
plot(fitdist(df_clean$log_body_mass, "norm"))

# normal with cube root-transformed body mass
plot(fitdist(df_clean$cube_body_mass, "norm"))
```

All of these plots show that the data (raw and transformed) are able to fit Gaussian and Gamma (for raw data) well. Will move on to fitting candidate models and then testing these.


## 5. Model formation
Generalised Linear Mixed (effect) Models (GLMMs) will be created here using the `glmmTMB` function. Multiple models will be formed based on different family and response parameters before undergoing intital residual and model testing, followed by statistical comparisons between models to select the most appropriate to take forward.

### A) Gaussian Model
The first model will be a GLMM with the Gaussian family (identity link), including all predictor variables to capture all potential effects of predictors on the response. Predictors can be dropped further down the line if necessary.  

Random effects will be included for park and observer as intercepts and not as random slopes as the effects of the predictors on the response is assumed to be the same across both sets of groups. Also with a small number of levels in each group (4 and 5 only), using random slopes could be unstable while also adding complexity to the model. 


```{r, error=FALSE, message=FALSE}
# glmm using glmmTMB
glmm_gauss <- glmmTMB(

  # fixed effects:
  body_mass ~ landscape +
    vegetation +
    lion_presence +
    distance_to_water_km +
    average_annual_rainfall_mm +
    predator_density_per_sq_km + 
    
    # park as random effect - each gets own intercept
    (1 | park) +
    # observer as random effect - each gets own intercept
    (1 | observer),
  
  data = df_clean,
  
  # using gaussian family
  family = gaussian()
)
```

Warning error produced upon creation:
```
Warning in finalizeTMB(TMBStruc, obj, fit, h, data.tmb.old) :
Model convergence problem; non-positive-definite Hessian matrix. See vignette('troubleshooting')
```
This error suggests an error in computing standard errors, meaning this model is essentially untrustworthy. Will also check for perfect collinearity below in case this is the reason. 

```{r}
# test if any predictors are perfectly collinear in this gauss model
matrix_gauss <- model.matrix(glmm_gauss)
findLinearCombos(matrix_gauss)
```

No perfectly collinear predictors found, error may be due to an overly complex model. However, residual and model checks can still take place.

A. Gauss model assessment
A.1 Check model summary:
 
```{r}
 summary(glmm_gauss)
```

This model summary contains no standard errors. `park` and `observer` as random effects have very low variances here, so the model may be treating them as fixed effects which can cause errors. Will continue with next checks and build further models.

A.2 Residual checks with `DHARMa`:
`DHARMa` tests simulate residuals from models by simulating repeated datasets using the model. It then tests where the observed data fall in thissimulated distribution. If the model is a good fit the residuals should follow a uniform pattern, and if not the model may be misfit and unsuitable.

```{r}
glmm_gauss_check <- simulateResiduals(fittedModel = glmm_gauss)
plot(glmm_gauss_check)
```
 
These simulates residuals from the GLMM show no deviation from the expected distribution indicating that the model assumptions were well met.

A.3 Multicollinearity check

```{r}
# check_collinearity(glmm_gauss)
```

This `check_collinearity()` did not work ("data frame is empty") -> likely cause is the Hessian error from during the build. Again, checks will continue.

A.4 Overdispersion checks
Testing whether a greater variability of data is seen when using the model compared to the actual observed data:

```{r}
check_overdispersion(glmm_gauss)
```

None detected

A.5 `performance` model check
`check_model` tests a fitted model for normality of residuals, normality of random effects, linear relationship, homogeneity of variance, and multicollinearity:

```{r, fig.width=10, fig.height=10}
# run check model
# check_model(glmm_gauss)
```

This also did not work, as such `glmm_gauss` is not suitable as it stands to be a candidate model, and further models will be made. The main cause is probably due to very low variance values from the random effects (scale of e-17 and e-11), so the model essentially treats them as fixed which can cause Hessian errors when calculating standard errors.


### B) Gamma Model (second GLMM)
Will also form a GLMM with a Gamma family and log link, again using all predictor variables.

```{r}
glmm_gamma <- glmmTMB(
  
  body_mass ~ landscape +
    vegetation +
    lion_presence +
    distance_to_water_km +
    average_annual_rainfall_mm +
    predator_density_per_sq_km +  
    
    # park as random effect - each gets own intercept
    (1 | park) +
    # observer as random effect - each gets own intercept
    (1 | observer),
  
  data = df_clean,
  
  # gamma with log link
  family = Gamma(link = "log")
)
```

B. Gamma model assessment
B.1 Check model summary:
 
```{r}
 summary(glmm_gamma)
```

Much larger variances for the random effects compared to `glmm_gauss`, and standard errors calculated.

B.2 Residual checks with DHARMa:

```{r}
glmm_gamma_check <- simulateResiduals(fittedModel = glmm_gamma)
plot(glmm_gamma_check)
```
 
Again, these simulates residuals from the GLMM show no deviation from the expected distribution indicating that the model assumptions were well met.

B.3 Collinearity check

```{r}
check_collinearity(glmm_gamma)
```

Low correlation for multicollinearity across these predictors which is good.

B.4 Overdispersion checks

```{r}
check_overdispersion(glmm_gamma)
```

None detected which is good.

B.5 `performance` model check

```{r, fig.width=10, fig.height=10}
# run check model
check_model(glmm_gamma)
```

### C) Log-transformed Gaussian Model (third GLMM)
Based on earlier data distribution checks and normality checking, and to try to mitigate the error from the first gaussian model, a log(response)-transformed gaussian GLMM will also be formed including all predictors:

```{r}
glmm_gauss2 <- glmmTMB(
  
  # log transformed response 
  log(body_mass) ~ landscape +
    vegetation +
    lion_presence +
    distance_to_water_km +
    average_annual_rainfall_mm +
    predator_density_per_sq_km +  
    
    # park as random effect - each gets own intercept
    (1 | park) +
    # observer as random effect - each gets own intercept
    (1 | observer),
  
  data = df_clean,
  family = gaussian()
)
```

C. Gauss2 model assessment
C.1 Check model summary:
 
```{r}
 summary(glmm_gauss2)
```

C.2 Residual checks with DHARMa:

```{r}
glmm_gauss2_check <- simulateResiduals(fittedModel = glmm_gauss2)
plot(glmm_gauss2_check)
```
 
Again, these simulates residuals from the GLMM show no deviation from the expected distribution indicating that the model assumptions were well met.

C.3 Collinearity check

```{r}
check_collinearity(glmm_gauss2)
```

C.4 Overdispersion checks

```{r}
check_overdispersion(glmm_gauss2)
```

None detected

C.5 `performance` model check

```{r, fig.width=10, fig.height=10}
# run check model
check_model(glmm_gauss2)
```


### D) Cube root-transformed Gaussian Model (fourth GLMM)
Based on earlier data distribution checks and normality checking, a cube root-transformed response gaussian GLMM will also be formed including all predictors:

```{r}
glmm_gauss3 <- glmmTMB(
  
  # cube root-transformed response 
    cube_body_mass ~ landscape +
    vegetation +
    lion_presence +
    distance_to_water_km +
    average_annual_rainfall_mm +
    predator_density_per_sq_km +  
    
    # park as random effect - each gets own intercept
    (1 | park) +
    # observer as random effect - each gets own intercept
    (1 | observer),
  
  data = df_clean,
  
  family = gaussian()
)
```

D. Gauss model assessment
D.1 Check model summary:
 
```{r}
 summary(glmm_gauss3)
```

D.2 Residual checks with DHARMa:

```{r}
glmm_gauss3_check <- simulateResiduals(fittedModel = glmm_gauss3)
plot(glmm_gauss3_check)
```

Again, these simulates residuals from the GLMM show no deviation from the expected distribution indicating that the model assumptions were well met.

D.3 Collinearity check

```{r}
check_collinearity(glmm_gauss3)
```

D.4 Overdispersion checks

```{r}
check_overdispersion(glmm_gauss3)
```

None detected

D.5 `performance` model check

```{r, fig.width=10, fig.height=10}
# run check model
check_model(glmm_gauss3)
```


## 6. Model testing
As there are currently four models which have all passed preliminary `DHARMa` residual and `check_model()` checks, they can be pitted against each other directly using `AIC` (Akaike information criterion) and `BIC` (Bayesian information criterion) tests.  

These two tests try to balance how well a model fits with its complexiy (models with more parameters are penalised harder to avoid overfitting). A lower score indicates a batter model in both AIC and BIC, with BIC penaising complexity more than AIC.  

AIC:

```{r}
AIC(glmm_gauss, glmm_gamma, glmm_gauss2, glmm_gauss3)
```

BIC:

```{r}
BIC(glmm_gauss, glmm_gamma, glmm_gauss2, glmm_gauss3)
```

In both AIC and BIC tests `glmm_gauss2` scored the lowest of all models (`glmm_gauss` received N/A, likely due to inability in computing standard errors when it was made). From this, `glmm_gauss2` will be taken forward as the candidate model.

A visualistaion of the above tests with plotting of observed mass vs predicted mass from each model, using `modelbased`:

```{r, message=FALSE, error=FALSE}
# get predicted data frame for each model based on the observed predictor values at each observation using estimate_expectation()
pred_gauss <- estimate_expectation(glmm_gauss)
pred_gamma <- estimate_expectation(glmm_gamma)
pred_gauss2 <- estimate_expectation(glmm_gauss2)
pred_gauss3 <- estimate_expectation(glmm_gauss3)

# create df of predicted values based off models' body mass (stored as 'Predicted')
models_predictions <- tibble(
  observed = df_clean$body_mass,
  predicted_gamma = pred_gamma$Predicted,
  predicted_gauss = pred_gauss$Predicted,
  # back transform to fit on correct axis:
  # exponentiate predicted
  predicted_gauss2 = exp(pred_gauss2$Predicted),
  # cube predicted
  predicted_gauss3 = (pred_gauss3$Predicted)^3
)

# create observed vs. predicted plot for all models:
models_predictions %>%
  
  ggplot() +
  # add points from each model including gauss2 model
  geom_smooth(aes(x = observed, y = predicted_gamma, colour = "gamma"), method = "lm", se = FALSE) +
  geom_smooth(aes(x = observed, y = predicted_gauss, colour = "gauss"), method = "lm", se = FALSE) +
  geom_smooth(aes(x = observed, y = predicted_gauss2, colour = "gauss2"), method = "lm", se = FALSE) +
  geom_smooth(aes(x = observed, y = predicted_gauss3, colour = "gauss3") , method = "lm", se = FALSE) +
  
  # 1:1 observed body mass from df_clean
  geom_line(aes(x = observed, y = observed), linetype = "dashed") +
  
  # adding a legend 
  scale_colour_manual(name = NULL,
                     breaks = c("gamma", "gauss", "gauss2", "gauss3"),
                     labels = c("Gamma Model", "Gauss Model", "Gauss2 Model", "Gauss3 Model"),
                     values = c("#1ABC9C", "#154360", "#FF5733", "#FFC300")) +
  
  ggtitle("Observed vs Predicted body mass for each model") +
  ylab("Body mass (predicted)") +
  xlab("Body mass (observed)") +
  theme_lucid()

```

As the dashed line is a 1:1 observed set of values, clearly all of the model plots are over fitting at lower masses, and under fitting at higher masses, despite passing residual and other model checks earlier. Attempts will be made to refine the models further to better predict observed values if possible while not overfitting. 

## 7. Model refining
With `glmm_gauss2` selected as the candidate model, model parameters will next be refined and further testing will take place, along with visualisation, in order to determine the most representative model for these data.

Another summary of `glmm_gauss2`:
```{r}
summary(glmm_gauss2)
```

### 7.1 Model simplifying
From the summary above I can see that only `distance_to_water_km` is a significant predictor in giraffe body mass in this model, and that there are two very non-significant predictors: `preditor_density_per_sq_km` (0.956) and `lion_presencePresent` (0.452). Both of these are clearly related too, and a model will be made without them to test if it may be simplified:

```{r}
glmm_gauss4 <- glmmTMB(
  
  # log transformed response with any predator predictors removed
  log(body_mass) ~ landscape +
    vegetation +
    distance_to_water_km +
    average_annual_rainfall_mm +

    # park as random effect - each gets own intercept
    (1 | park) +
    # observer as random effect - each gets own intercept
    (1 | observer),
  
  data = df_clean,
  family = gaussian()
)
```

Check model summary:
 
```{r}
 summary(glmm_gauss4)
```


Check model residuals with DHARMa:  

```{r}
glmm_gauss4_check <- simulateResiduals(fittedModel = glmm_gauss4)
plot(glmm_gauss4_check)
```
```{r, fig.width=10, fig.height=10}
check_model(glmm_gauss4)
```

`check_model` and DHARMa residual tests all passed, so on to AIC and BIC tests:

```{r}
AIC(glmm_gauss, glmm_gamma, glmm_gauss2, glmm_gauss3, glmm_gauss4)
```

```{r}
BIC(glmm_gauss, glmm_gamma, glmm_gauss2, glmm_gauss3, glmm_gauss4)
```

Here `glmm_gauss4` scored lower than all other models, as well as being simpler (8 df vs 10). This can be considered an improved model over `glmm_gauss2`, placing it as the current candidate model.  

Other models were also testes but cannot be shown due to space constraints. Models with predictor interactions and random effects were included for:  

- Water distance and vegetation interaction  
- Rainfall and landscape interaction    
- Rainfall and vegetation interaction   
- Distance to water and landscape interaction    
- Distance to water across parks as a random slope was also included  

All of these altered models failed to improve on `glmm_gauss4` in terms of AIC/BIC scores while also adding complexity. As such the `glmm_gauss4` model will remain and be used as the final model.


## 8. Model interpretation
### 8.1 Model equation

Summary of `glmm_gauss4`:

```{r}
summary(glmm_gauss4)
```

Therefore `glmm_gauss4` can be expressed as:  
$$
\text{log(Body Mass}_{i}\text{)} = \beta_0 + \beta_1\text{Landscape}_{i} + \beta_2\text{Vegetation}_{i} + \beta_3\text{Distance to Water}_{i} + \beta_4\text{Rainfall}_{i} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}
$$

Where:  

- $i$ represents the observation
- $\beta_0$ is the expected body mass when all other predictors are at reference level or mean values.  
- $\beta$ values are the model variable coefficients  
- $\text{Landscape}_{i}$ -> BOOL landscape value: hilly (=1) or flat (=0)  
- $\text{Vegetation}_{i}$ -> BOOL vegetation value: open (=1) or dense (=0)  
- $\text{Distance to Water}_{i}$ is the distance to water (km) 
- $\text{Rainfall}_{i}$ is the average annual rainfall (mm) 
- $b_{park[i]}$ is the intercept shift for park in observation $i$  
- $b_{observer[i]}$ is the intercept shift for observer in observation $i$  
- $\epsilon_{i}$ is the error term (for residual error)  

With coefficients: 
$$
\text{log(Body Mass}_{i}\text{)} = 7.15 - 0.040\cdot\text{Landscape}_{i}\text{} - 0.038\cdot\text{Vegetation}_{i} + 0.011\cdot\text{Distance to Water}_{i}\text{(km)} - (6.16\times 10^{-5})\cdot\text{Annual Rainfall}_{i}\text{(mm)} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}
$$

And back transformed:  
$$
\text{Body Mass}_{i} = e^{7.15 - 0.040\cdot\text{Landscape}_{i}\text{} - 0.038\cdot\text{Vegetation}_{i} + 0.011\cdot\text{Distance to Water}_{i}\text{(km)} - (6.16\times 10^{-5})\cdot\text{Annual Rainfall}_{i}\text{(mm)} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}}
$$

### 8.2 Significant predictor: distance from water
This model suggests that the only significant predictor on giraffe body mass is the Distance from water ($p = 0.032$), and its effect can be shown using the model as: 

$$
\text{Body Mass}_i \propto e^{0.011\cdot\text{Distance to Water}_{i}\text{(km)}}
$$

$$
e^{0.011} \approx 1.011
$$
$$
\therefore \text{Body Mass} \propto 1.011^{\text{Distance to Water (km)}}
$$

So the model predicts that giraffe mass increases by 1.1% for every km further from water the giraffe is when observed, all other predictors constant. 

## 9. Model fit

This model predictions can be plotted against the observed values to visualise how well it fits the observed data:

```{r}
# create predicted and observed data frame to get predicted values for mass from the model based on the predictor values at each true observation:
fit_model_gauss4 <- data.frame(
  
  # back transform from log to get predicted on the correct scale:
  "Predicted" = exp(predict(glmm_gauss4, type = "response")),
  "Observed" = df_clean$body_mass
  
)

# plot predicted vs observed with 1:1 (observed:observed) comparison line
ggplot(fit_model_gauss4) +
  
  # each point (x) = observed, (y) = model prediction
  geom_point(aes(x = Observed, y = Predicted,)) +
  
  # add linear regression of observed:predicted
  geom_smooth(
    aes(x = Observed, y = Predicted, colour = "gauss4"),
    method = lm,
    se = FALSE) +
  
  # add 1:1 linear regression to plot to check main regression
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", aes(colour = "1:1")) +

  # aesthetics
  scale_colour_manual(name = NULL,
                      breaks = c("1:1", "gauss4"),
                      labels = c("Observed:Observed", "Gauss4 Model"),
                      values = c("black", "#FF5733", "red")) +
  
  xlab("Body mass (observed)") +
  ylab("Body mass (predicted)") +
  
  ggtitle("Predicted vs. Observed Giraffe Body Mass for Gauss4 Model") +
  theme_lucid()

```

Points represent the predicted mass value from the `glmm_gauss4` model based on the parameters from the true observations in the data. The fitted line shows that the model tends to over predicts lower body masses, and under predict higher body masses compared to the observed values.

Variance in the observed response (log(body mass)) captured by the model vs. residual error can be quantified using `r2_nakagawa()` from `performance`:

```{r}
r2_nakagawa(glmm_gauss4)
```

From this, the conditional r-squared (random and fixed effects in the model) only capture 22.8% of the variance, with marginal r-squared (fixed effects only) capturing 3.3% of the variance. Therefore the random effects account for about 19.5% of the variance in the response.  This means that the remaining 77.2% of the variance cannot be explained by the model whicih could be caused by differences not accounted for by the chosen predictors, or other biological factors not included in the model. 

## 10. Visual interpretations

To visualise and interpret the model clearly, marginal means and effects will be used to interpret how each predictor affects giraffe bosy mass using the model. The package `modelbased` will be used for this.

Marginal means will be used for `landscape` and `vegetation`, while marginal effects will be calculated for the continuous predictors. 








