---
title: "BIOLM0039 - Investigating Giraffe Body Mass"
author: "Archie Benn (1935263)"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Set workspace
WORDCOUNTER: `pandoc BIOLM0039_exam.Rmd --to=plain | wc -w`


```{r, echo = TRUE, message=FALSE, error=FALSE}
# reset environment
rm(list=ls())

# load libraries
library(tidyverse)
library(wesanderson)
library(performance)
library(Hmisc)
library(ggpubr)
library(rstatix)
library(fitdistrplus)
library(MASS)
library(see)
library(TMB)
library(glmmTMB)
library(DHARMa)
library(data.table)
library(caret)
library(modelbased)
library(broom.mixed)
library(dotwhisker)
library(ggeffects)
```

## 2. Import and check Data
Load in the data (tsv file saved from GitHub) as a tibble:

```{r, message=FALSE}
df <- read_tsv("../data/mass_sj19031.tsv")
```

Describe to get an overview of the data:

```{r}
describe(df)
```

`df` contains the following:  
- Giraffe mass (kg) collected from 5 observers across 4 parks  
- Two level categorical data for `lion_presence`, `vegetation`, and `landscape`    
- Continuous data on average annual rainfall, distance to water, and predator density  
- Missing values in `landscape`, `vegetation`, `lion_presence`, `env_value`  
- 5 distinct values for `lion_presence` -> Typos in 'Absent' or 'Present' 

## 2. Tidying data
### 2.1 Outliers
Boxplot to check outliers and see a general distribution of body mass over the full dataset `df`:

```{r}
# boxplot of entire dataset mass to check potential outliers
mass_box <- df %>%
  
  ggplot(aes(x = "", y = body_mass)) +
  
  # reduce width to 0.25
  geom_boxplot(width = 0.25, fill = wes_palette("IsleofDogs1")[1]) +
  
  # aesthetics
  theme_lucid() +
  labs(title = "Box plot of giraffe mass across original dataset") +
  
  # centre title
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") +
  ylab("Body mass (kg)")

mass_box
```

Box plot identifies 5 potential mass outliers in `df` using the IQR criterion.    

An outlier data frame will be created using which will be used to remove these data during cleaning. This is justified biologically as giraffes rarely exceed 2000kg and outliers are identified statistically:

```{r}
# create outlier data frame
outlier_df <- identify_outliers(data = df,
                  variable = "body_mass")
```

The dataset `df` is also in long format, and to have each row as an individual observation will be transformed to wide format during cleaning.

### 2.2 Create cleaned data frame
A clean data frame will be created based on the information above:

```{r, message=FALSE, error=FALSE}
# using tidyverse
df_clean <- df %>%
  
  # remove outliers using the outlier data frame:
  anti_join(outlier_df)  %>%
  
  # long to wide format df:
  pivot_wider(names_from = env_var, values_from = env_value) %>%

  # remove NA values including those in the new 'wide columns':
  drop_na(landscape, vegetation, lion_presence, distance_to_water_km, average_annual_rainfall_mm, predator_density_per_sq_km) 


# fix typos in lion_presence (no need to remove as very clearly 'Present' or 'Absent' typos)
df_clean$lion_presence[df_clean$lion_presence == "Preesnt"] <- "Present" 
df_clean$lion_presence[df_clean$lion_presence == "Preseet"] <- "Present"
df_clean$lion_presence[df_clean$lion_presence == "Asbent"] <- "Absent"
```

`df_clean` now has no missing variables, fixed `lion_presence' typos, no statistical outiers based on the original `df`, and each row is an individual observation.

## 3. Variables and hypotheses
### 3.1 Variables
`df_clean` contains variables as follows:  

**Response:**  
- `body_mass`  

**Predictors (fixed):**  
 - Categorical = `landscape`, `vegetation`, `lion_presence`  
 - Continuous = `distance_to_water_km`, `average_annual_rainfall_mm`, and `predator_density_per_sq_km`  
 
**Predictors (random):**  
These two groups only have 4 and 5 levels each which could make variance estimates unstable but will be included as random effects to account for clustering among observations.
 - Groups/grouping: `park`, `observer`   
 
These data are spatially structured, and no information on temporal structure is provided. Need to consider if random effects will be considered as slopes or intercepts, and any interactions when forming models.  

The response (body mass) is continuous and $\geq$ 0, which allows for Gaussian and/or Gamma if a GLMM is to be used.

#### 3.1.1 Crossed or nested random effects?
If any observer makes observations in $\gt$ 1 park, the random effects should be crossed, but if observers are contained within only one park and do not make observations elsewhere, they can be treated as nested. A contingency table can show the number of observations for each observer across each park:

```{r}
table(df_clean$observer, df_clean$park)
```

So each observer makes observations at all parks, therefore the random effects will be crossed.  

### 3.2 Hypotheses
The following null and alternative hypotheses are proposed to test giraffe body mass:

$$
H_0: \text{None of the factors included in this dataset have a significant effect on giraffe body mass}\\
H_1: \text{At least one factor included in this dataset has a significant effect on giraffe body mass}
$$

## 4. Raw data exploration
### 4.1 Giraffe mass across each park and observer:
#### Park violin plot:

```{r}
plot_mass_parks <- df_clean %>%
  
  ggplot(mapping = aes(x = park, y = body_mass)) +
  
  # violin plot 
  geom_violin(aes(fill = park)) + 
  
  # hide redundant legend
  guides(fill = "none") +
  
  # add boxplot 
  geom_boxplot(width = 0.1, col = "black")+
  
  
  
  # aesthetics
  scale_fill_manual(values = wes_palette("Darjeeling2")) +
  ggtitle("Violin plot of giraffe mass across the parks") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45)) +
  ylab("Body mass (kg)") + 
  xlab("Park") 

plot_mass_parks
```


Zebra Horizon Park appears to have a slightly raised mean giraffe mass compared to the other three parks.

#### Observer violin plot  

```{r}
df_clean %>%
  
  ggplot(mapping = aes(x = observer, y = body_mass)) +
  
  # violin plot 
  geom_violin(aes(fill = observer)) + 
  
  # hide redundant legend
  guides(fill = "none") +
  
  # add boxplot 
  geom_boxplot(width = 0.1, col = "black")+
  
  
  
  # aesthetics
  scale_fill_manual(values = wes_palette("AsteroidCity1")) +
  ggtitle("Violin Plot of Giraffe Mass Recorded Across the Observers") +
  theme_minimal() +
  ylab("Body mass (kg)") + 
  xlab("Observer") 

```

### 4.2 Scatter plots
Log distance to water:

```{r}
# scatter plot
df_clean %>%
  
  # log the distance to water (+1 for positive x axis)
  ggplot(aes(x=log(distance_to_water_km + 1), y=body_mass)) +
  
  # plot data points
  geom_point(col=wes_palette("AsteroidCity2")[1]) +
  
  # add linear regression
  geom_smooth(method = "lm", se = FALSE, color = "black", formula = y~x, linetype = "dashed") +
  
  # aesthetics
  ggtitle("Giraffe Mass Against Distance to Water Across Parks") +
  xlab("log(Distance to water (km))") +
  ylab("Body mass (kg)") +
  
  # group by park
  facet_wrap(. ~ park) +
  theme_lucid() 
```

The distance to water appears to have a slight positive effect on body mass across 3 out of the 4 parks.

Similar plots were created for average annual rainfall and predator density, both showing weak relationships to body mass.

### 4.3 Distribution and transformation checks
Body mass is strictly positive, as such the main families involved for a model are likely Gaussian and Gamma. A histogram of body mass and Box-Cox will be used to assess potential transformations to attempt later.

#### Histogram with overlaid reference normal distribution:

```{r, message=FALSE, error=FALSE, warning=FALSE}
df_clean %>%
  
  ggplot(aes(x = body_mass)) +
  
  # add histogram with bin width at 50kg and scale using ..density..
  geom_histogram(aes(y = ..density..), fill = wes_palette("Cavalcanti1")[4], color = "black", binwidth = 50) +
  
  # overlay normal distribution with same mean and sd as body mass
  stat_function(fun = dnorm,
                args = list(
                  mean = mean(df_clean$body_mass),
                  sd = sd(df_clean$body_mass)),
                linetype = "dashed") +
  
  # aesthetics
  theme_minimal() +
  labs(title = "Histogram of giraffe body mass") +
  xlab("Body mass (kg)") +
  ylab("Density") 
```

Histogram indicates a right skewed body mass in this cleaned data. Transformations may aid approximating to normal.

#### Box-Cox
Box-Cox from `MASS` can help to see whether a power transformation will approximate the data to normal for Gaussian models. Box-Cox provides a value, $\lambda$, which has the greatest log likelihood for approximating normal given the data. 

```{r}
# first create a linear model of the body mass, intercept only
lm_body_mass  <- lm(body_mass ~ 1, df_clean)

# apply box-cos to generate a log-likelihood plot which tests lambda from -2 to 2, step 0.1
boxcox(lm_body_mass, lambda = seq(-2, 2, 0.1))
```

From this, highest log-likelihood $\lambda \approx$ 0.25 which is between a log ($\lambda = 0$) and a cube root ($\lambda = 1/3$) transformation, and means non-transformed ($\lambda = 1$) is unlikely to fit as well. This, along with the positive and right skewed response, suggests the following transformations for `fitdistr` testing below:  

- Gamma family with raw data  
- Log and cube-root transformed responses with Gaussian family

### 4.4 Family checks
Using `fitdistr` to assess potential error structures:

```{r}
# log response with gaussian 
plot(fitdist(log(df_clean$body_mass), "norm"))
```

Above plot demonstrates a suitable fit for log transformed response with Gaussian, and was repeated for both Gamma and cube-root transformed Gaussian, with each showing adequate fit. Will move on to fitting candidate models.

## 5. Model fitting
### Log transformed Gaussian Model
Log transformed response GLMM with the Gaussian family (identity link), including all predictors. Random effects will be included as intercepts (not slopes), as effects are assumed to be the same across levels within groups and to avoid overfitting with the limited number of levels in each (4 and 5).


```{r, error=FALSE, message=FALSE}
# fit random effects model
glmm_gauss_log <- glmmTMB(

  # log transformed response ~ all fixed effects:
  log(body_mass) ~ landscape +
    vegetation +
    lion_presence +
    distance_to_water_km +
    average_annual_rainfall_mm +
    predator_density_per_sq_km + 
    
    # crossed random effects (park and observer), each with its own intercept:
    (1 | park) +
    (1 | observer),
  
  data = df_clean,
  
  family = gaussian()
)

summary(glmm_gauss_log)
```

#### Model check 1: Residual structure with `DHARMa`:

```{r}
# simulate residuals for model with 1000 simulations (reduce stochastic error and stabilises values):
plot(simulateResiduals(glmm_gauss_log, n =1000))
```

No dispersion or deviation warnings indicate this model provides a good fit to the data.

#### Model check 2: Multicollinearity between predictors:

```{r}
check_collinearity(glmm_gauss_log)
```

VIF values ~1 indicate low multicollinearity between predictors in this model, with well estimated coefficients and $p$ values.

#### Model check 3: Overdispersion checks with `DHARMa`
For a more thorough overdispersion check:

```{r}
testDispersion(glmm_gauss_log)
```

Not significant ($p = 0.664$), no overdispertion detected.

#### Model check 4: `check_model` overview:

```{r, fig.width=10, fig.height=10}
check_model(glmm_gauss_log)
```

All diagnostics indicate this model suits the data well. However, the residuals variance does seem to display some heteroscedasticity, with lower residuals at higher fitted values. This pattern may be due to missing predictors which could affect higher body masses disproportionately.

### Fitting two further models

```{r, message=FALSE, error=FALSE, warning=FALSE}
# create gamma model (without log transformed response)
glmm_gamma <- update(glmm_gauss_log, formula = body_mass ~ ., family = Gamma(link = "log"))
# cube-root response transformed model
glmm_gauss_cubed <- update(glmm_gauss_log, formula = (body_mass)^(1/3) ~ .)
```

These two models also underwent the same checks as the log Gaussian model above, and both fit well, with diagnostics performance comparable to `glmm_gauss_log`.

## 6. Model testing
### 6.1 Model predictions visualised
A visualisation of the above tests with plotting of observed mass vs predicted mass from each model, using `modelbased`:

```{r, message=FALSE, error=FALSE}
# get predicted data frame for each model based on the observed predictor values at each observation using estimate_expectation()
pred_gamma <- estimate_expectation(glmm_gamma)
pred_gauss_log <- estimate_expectation(glmm_gauss_log)
pred_gauss_cubed <- estimate_expectation(glmm_gauss_cubed)

# create df of predicted values based off models' body mass (stored as 'Predicted')
models_predictions <- tibble(
  observed = df_clean$body_mass,
  predicted_gamma = pred_gamma$Predicted,
  
  # back transform predictions from log and cubed models:
  predicted_gauss_log = exp(pred_gauss_log$Predicted),
  predicted_gauss_cubed = (pred_gauss_cubed$Predicted)^3
)

# create observed vs. predicted plot for all models:
models_predictions %>%
  
  ggplot() +
  
  # add loess line from each model's predictions
  geom_smooth(aes(x = observed, y = predicted_gamma, colour = "gamma"), se = FALSE) +
  geom_smooth(aes(x = observed, y = predicted_gauss_log, colour = "gauss_log"), se = FALSE) +
  geom_smooth(aes(x = observed, y = predicted_gauss_cubed, colour = "gauss_cubed"), se = FALSE) +
  
  # add slope = 1 line for observed = predicted
  geom_abline(aes(intercept = 0, slope = 1, colour = "1:1"),  linetype = "dashed",) +  
  
  # adding a legend 
  scale_colour_manual(name = NULL,
                     breaks = c("gamma", "gauss_log", "gauss_cubed", "1:1"),
                     labels = c("Gamma Model", "Log Gauss Model", "Cube-root Gauss Model", "Predicted = Observed"),
                     values = c(wes_palette("AsteroidCity1"))) +
  
  ggtitle("Observed against Predicted Body Mass Across Models") +
  ylab("Body mass (kg) - predicted") +
  xlab("Body mass (kg) - observed") +
  theme_lucid()

```

All models display over predicting at lower masses and under predicting at higher masses, indicating an inability to adequately capture the body mass variance towards extremes. This reflects the heteroscedasticity mentioned earlier which can pull the predicted values at high body masses back towards the mean. Predictions towards the extremes should be taken with caution.  

### 6.2 `ANOVA` model comparisons
As there are currently three working models which have all passed `DHARMa` residual and `check_model()` checks and appear to predict very similar body masses, their fits can be directly quanitified against each other using `anova()` with `AIC` and `BIC` acting as comparative scores. 

AIC and BIC try to balance how well a model fits with its complexity (models with more parameters are penalised harder to avoid overfitting. A lower score indicates a batter model in both AIC and BIC, with BIC penalising complexity more than AIC:  

```{r}
anova(glmm_gauss_log, glmm_gamma, glmm_gauss_cubed)
```

Despite all three models performing well in diagnostic tests,  AIC and BIC results show `glmm_gauss_log` scored the lowest in both and as such will be taken forward as the candidate model. 

## 7. Model refining
With `glmm_gauss_log` selected as the candidate model, model parameters will next be refined and further testing will take place, along with visualisation, to further refine the model if possible.

Another summary of `glmm_gauss2`:
```{r}
summary(glmm_gauss_log)
```


### 7.1 Model simplifying
From the summary above only `distance_to_water_km` is a significant predictor in giraffe body mass in this model, and there are two non-significant predictors: `predator_density_per_sq_km` ($p = 0.956$) and `lion_presencePresent` ($p = 0.452$). Also, the coefficient for average rainfall is so small ($\approx$ 5.7e-5) it's negligible in this model and will be removed. 

A model will be made without these three predictors to test if reducing complexity still produces a valid model:

```{r}
# update candidate model to remove the lion/predator and rainfall predictors:
glmm_gauss2_log <- update(glmm_gauss_log, formula. = . ~ . - average_annual_rainfall_mm - lion_presence - predator_density_per_sq_km)

# check model summary:
 summary(glmm_gauss2_log)
```

Look at tidy summary of effects:

```{r}
broom.mixed::tidy(glmm_gauss2_log, conf.int = TRUE)
```

Comparing `glmm_gauss_log` and `glmm_gauss2_log` with ANOVA:

```{r}
anova(glmm_gauss_log, glmm_gauss2_log)
```

Removing the above fixed effects did not significantly worsen the model fit ($p = 0.456$), and the lower AIC/BIC vales mean this model is preferred for parsimony, even at the slightly reduced 'biological completeness' of the model from removing predictors.

Check model residuals with DHARMa, and `check_model()`:  

```{r}
# simulate residuals with 1000 simulations:
glmm_gauss2_log_res <- simulateResiduals(glmm_gauss2_log, n = 1000)
plot(glmm_gauss2_log_res)
```
```{r, fig.width=10, fig.height=10}
check_model(glmm_gauss2_log)
```

Still passes residual testing and other diagnostics. As such the `glmm_gauss2_log` model will be taking forward as the final model.

## 8. Model fit and interpretation
This model predictions can be plotted against the observed values to visualise how well it fits the observed data:

```{r, message=FALSE, error=FALSE, warning=FALSE}
# create predicted and observed data frame to get predicted values for mass from the model based on the predictor values at each true observation:
fit_model_gauss2_log <- data.frame(
  
  # back transform from log to get predicted on the correct scale:
  "Predicted" = exp(predict(glmm_gauss2_log, type = "response")),
  "Observed" = df_clean$body_mass
  
)

# plot predicted vs observed with 1:1 (observed:observed) comparison line
ggplot(fit_model_gauss2_log) +
  
  # each point (x) = observed, (y) = model prediction
  geom_point(aes(x = Observed, y = Predicted,)) +
  
  # add loess line of observed:predicted from the model
  geom_smooth(
    aes(x = Observed, y = Predicted, colour = "gauss"),
    method = loess,
    se = FALSE) +
  
  # add slope = 1 line for observed = predicted
  geom_abline(aes(intercept = 0, slope = 1, colour = "1:1"),  linetype = "dashed",) +

  # aesthetics
  scale_colour_manual(name = NULL,
                      breaks = c("1:1", "gauss"),
                      labels = c("Observed = Predicted", "Log Gauss2 Model"),
                      values = c("black", wes_palette("AsteroidCity3")[1])) +
  
  xlab("Body mass (observed)") +
  ylab("Body mass (predicted)") +
  
  ggtitle("Predicted vs. Observed Giraffe Body Mass for Log Gauss2 Model") +
  theme_lucid()

```

Points represent the predicted mass value from the `glmm_gauss2_log` model based on the predictor values from the true observations in the data. The fitted loess line shows that the model tends to over predicts lower body masses, and under predict higher body masses compared to the observed values which could be due to missing predictors in the data as mentioned above, however overall residuals and model checks indicate a good fit. As the main aim is to test any significant predictors on body mass, this model is appropriate.

Variance in the observed response (log(body mass)) captured by the model vs. residual error can be quantified using `r2_nakagawa()` from `performance`:

```{r}
r2_nakagawa(glmm_gauss2_log)
```

From this, the conditional r-squared (random and fixed effects in the model) only capture 22.1% of the variance, with marginal r-squared (fixed effects only) capturing 2.7% of the variance. Therefore the random effects account for about 19.5% of the variance in the response.  

This means that the remaining 77.9% of the variance cannot be explained by the model which could be caused by differences not accounted for by the chosen predictors, or other biological factors not included in the model. This is quite a large amount of variance to be unexplained by the model, and is possibly due to missing other predictors in the data. Missing a significant amount of data which could impact giraffe body mass reduces a model's prediction capabilities, and therefore decreases the amount of variance in the response which can be explained by the model.

## 9. Model equation and predictor significance
Exact coefficients are not strictly necessary as the main aim is to assess which factors significantly impact giraffe body mass, but it can be helpful to explicitly understand the model in log linear and back transformed forms.

### 9.1 Model equation
Summary of `glmm_gauss2_log`:

```{r}
summary(glmm_gauss2_log)
```

From this model summary `glmm_gauss2_log` can be expressed as:  
$$
\text{log(Body Mass}_{i}\text{(kg))} = \beta_0 + \beta_1\text{Landscape}_{i} + \beta_2\text{Vegetation}_{i} + \beta_3\text{Distance to Water}_{i} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}
\tag{1}
$$

Where:  

- $i$ represents the observation
- $\beta_0$ is the expected body mass when all other predictors are at reference level or mean values.  
- $\beta$ values are the model variable coefficients  
- $\text{Landscape}_{i}$ -> BOOL landscape value: hilly (=1) or flat (=0)  
- $\text{Vegetation}_{i}$ -> BOOL vegetation value: open (=1) or dense (=0)  
- $\text{Distance to Water}_{i}$ is the distance to water (km) 
- $b_{park[i]}$ is the intercept shift for park in observation $i$  
- $b_{observer[i]}$ is the intercept shift for observer in observation $i$  
- $\epsilon_{i}$ is the error term (the unexplained 77.9% of variance)  

With coefficients: 
$$
\text{log(Body Mass}_{i}\text{(kg))} = 7.09 - 0.041\cdot\text{Landscape}_{i}\text{} - 0.039\cdot\text{Vegetation}_{i} + 0.012\cdot\text{Distance to Water}_{i}\text{(km)} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}
$$

### 9.2 Back transformation 
Back transformation gets from log (natural) to the normal scale:  

$$
\text{Body Mass}_{i}\text{(kg)} = e^{7.09 - 0.041\cdot\text{Landscape}_{i}\text{} - 0.039\cdot\text{Vegetation}_{i} + 0.012\cdot\text{Distance to Water}_{i}\text{(km)} + b_{park[i]} + b_{observer[i]} + \epsilon_{i}}
$$

On this scale, the predictors now have multiplicative, rather than additive, effects on the body mass in this model. In other words a predictor with coefficient $\beta$ in the log scale now has a multiplicative effect of $e^{x\beta}$ on body mass for an $x$ unit increase in that continuous predictor (or $e^\beta$ vs. $1$ for the categorical predictors), all other effects constant.

### 9.3 Significant predictor: distance from water
This model suggests that the only significant predictor on giraffe body mass at the 5% significance level is the Distance from water ($p = 0.025$). The effect of this predictor in the model is: 

$$
\text{Body Mass}_i\text{(kg)} \propto e^{0.012\cdot\text{Distance to Water}_{i}\text{(km)}} \approx 1.012^{\text{Distance to Water (km)}}
$$


So the model predicts a small marginal effect: giraffe mass increases by $\approx$ 1.2% for every km further from water the giraffe is when observed, all other predictors constant, and at a significance of $p = 0.025$. 



## 10. Interpreting model fixed effects

```{r}
# data frame with fixed effect and confidence intervals:
model_fixed_effects_tidy <- broom.mixed::tidy(glmm_gauss2_log, conf.int = TRUE, conf.level = 0.95) %>%
  
  # don't include random effects (see next section):
  filter(effect == "fixed") %>%
  
  # remove intercept value
  filter(term!="(Intercept)")
  


# back transform estimates and confidences:
model_fixed_effects_tidy <- model_fixed_effects_tidy %>%
  
  # back transform
  mutate(
    estimate_bt = exp(estimate),
    conf.low_bt = exp(conf.low),
    conf.high_bt = exp(conf.high)
  )
```

```{r}
# plot fixed effect coefficients
ggplot(model_fixed_effects_tidy, aes(x=estimate, y=term)) + 
  
  # add fixed effect coefficient
  geom_point(shape = 21, fill = wes_palette("FantasticFox1")[2], size = 5)+
  
  # and 95% CIs
  geom_errorbar(aes(xmin=conf.low, xmax=conf.high),                       
              width=0.3, 
              col="grey50") +
  
  # red dashed line at x = 0
  geom_vline(xintercept =0, linetype = "dashed", col = "red") +
  
  # aesthetics
  xlab("Coefficient estimate") +
  ylab("Fixed effect") +
  ggtitle("Estimated Model Fixed Effect Coefficients at 95% CI") +
  theme_lucid() 

```

Visual of the statistically significant predictor against body mass:

```{r, message=FALSE, error=FALSE, warning=FALSE}
# retrieve back transormed predicted data frame using ggpredict at 95% CI
df_predict <- ggpredict(glmm_gauss2_log, terms = c("distance_to_water_km"), ci_level = 0.95, back_transform = TRUE)

# plot predictions for distance to water
plot(df_predict) +
  
  # overlay observed data
  geom_point(data = df_clean, aes(x = distance_to_water_km, y = body_mass), alpha = 0.5, color = wes_palette("Moonrise2")[1]) +
  
  # aesthetics
  xlab("Distance to Water (km)") +
  ylab("Body Mass (kg)") +
  ggtitle("Predicted Body Mass against Distance to Water with 95% CIs (other predictors constant)") +
  theme_lucid() + 
  geom_line(color = wes_palette("Moonrise2")[2], linewidth = 1.25) 

```

The points above are the observed data overlaid on the model's prediction for body mass against distance to water. This plot is an exponential on this normal scale with $\text{Body Mass} \propto 1.012^{\text{Distance to Water}}$ but approximates a linear plot over this range as the coefficient is so small (1.2% increase per unit distance).

## 11. Interpreting model random effects
From the above equation for `glmm_gauss2_log` (equation 1), there are random effects termed $b_{park[i]}$ and $b_{observer[i]}$. These are estimated multiplicative factors $e^{b_{park[i]}}$ and $e^{b_{observer[i]}}$ (on the original scale) from the expected population level body mass, and they represent deviations in body mass purely due to the level of the random effect, assuming fixed effects are constant.

To visualise the estimated multiplicative effect for each level, `ranef()` from `glmmTMB` will be used to extract random effects (park and observer):

### 11.1 Park

```{r}
# extract park random effects  with condVar attributes attached (as a list):
ranefs <- ranef(glmm_gauss2_log, condVar = TRUE)

# ranef contains the data frames within $cond, so to extract and mutate park df:
ranefs_park <- ranefs$cond$park %>%
  
  mutate(
    # make each row name a factored column for the data frames
    park = factor(rownames(.)),
    
    # back transform the intercept to get on the normal scale
    bt = exp(`(Intercept)`),
    
    # extract conditional variance from the 3d array it is stored in in this df:
    condVar = attr(., "condVar")[1, 1, ],
    
    # calculate (log) standard errors from condVar:
    se = sqrt(condVar),
  
    # back transform and add upper and lower bounds at 95% confidence intervals (1.96 x se)
    bt_lower = exp(`(Intercept)` - 1.96*se),
    bt_upper = exp(`(Intercept)` + 1.96*se)
  )

head(ranefs_park)
```

Now the data frame has the necessary estimated multiplicative factors (`bt`) for each level, they can be plotted along with the errors for each estimate. Note these factors are equivalent to $e^{b_{park[i]}}$ due to back transformation.

```{r, message=FALSE, error=FALSE}
# plot park random effect impact on body mass 

ggplot(ranefs_park, aes(x = bt, y = park)) +
  
  geom_point(col = "#F89C74", size = 5) + 
  
  # add effect horizontal error bars
  geom_errorbarh(aes(xmin = bt_lower, xmax = bt_upper), height = 0.25, col = "grey50") +
  
  # add vertical line for mean/expected across all parks
  geom_vline(xintercept = 1, col ="red3", linetype = "dashed") +
  
  ylab("Park") +
  xlab("Multiplicative factor on body mass ") +
  ggtitle("Estimated Impact of Park (Random Effect) on Body Mass at 95% CI") +
  
  theme_lucid()
```

Error bars here represent a 95% confidence interval (CI) after back transforming, and the red dashed line represents the population level expectation on the multiplicative scale after accounting for the fixed effects (x = 1 represents no effect on the multiplicative scale for body mass).

Park multiplicative factors, at 95% CI, range from ~0.88 to ~1.20, and indicate that some parks have a higher/lower baseline body mass. Confidence intervals indicate Zebra Horizon Park masses are ~10% higher than the expected population value (factor 1.10 [CI: 1.02 - 1.20]).

To assess whether park random effects contribute significantly to this model, a model without the parks will be fitted and then compared to `glmm_gauss2_log` with anova:

```{r}
# new model without park random effect:
glmm_no_park <- update(glmm_gauss2_log, ~ . - (1 | park))
# compare with anova:
anova(glmm_gauss2_log, glmm_no_park)
```

The above shows higher AIC/BIC values without the park, and a very significant chi squared value ($p = 2.21e-5$). This indicates meaningful body mass variation across parks and its inclusion improves model fit.


#### 11.2 Observer
And for the observer random effect, the same is done:  

```{r}
# and for observer (same as park above):
ranefs_observer <- ranefs$cond$observer %>%
  mutate(
    observer = factor((rownames(.))),
    bt = exp(`(Intercept)`),
    condVar = attr(., "condVar")[1, 1, ],
    se = sqrt(condVar),
    bt_lower = exp(`(Intercept)` - 1.96*se),
    bt_upper = exp(`(Intercept)` + 1.96*se)
  )



head(ranefs_observer)
```


```{r, message=FALSE, error=FALSE}
# plot observer random effect impact on body mass 
ggplot(ranefs_observer, aes(x = bt, y = observer)) +
  
  geom_point(col = "#154360", size = 5) + 
  
  # add effect horizontal error bars
  geom_errorbarh(aes(xmin = bt_lower, xmax = bt_upper), height = 0.25, col = "grey50") +
  
  # add vertical line for mean/expected across all observers
  geom_vline(xintercept = 1, col ="red3", linetype = "dashed") +
  
  ylab("Observer") +
  xlab("Multiplicative factor on body mass") +
  ggtitle("Estimated Impact of Observer (Random Effect) on Body Mass at 95% CI") +
  
  theme_lucid()
```

Error bars here represent a 95% CI after back transforming, and the red dashed line represents the population level expectation on the multiplicative scale after accounting for the fixed effects (x = 1 represents no effect on the multiplicative scale for body mass).

Multiplicative factors, at 95% CI, range from ~0.83 to ~1.20 across observers. The 95% confidence intervals of observers 1 and 2 do not cross the x = 1 line, indicating that observer 1 records masses ~10% higher (factor 1.10 [CI 1.01 - 1.20]), and observer 2 ~9% lower (factor 0.91 [CI 0.84 - 0.99]), than the population expectation in this model.

To assess whether observer random effects contribute significantly to this model, a model without the observers will be fitted, as above with parks:

```{r}
# new model without observer random effect:
glmm_no_observer <- update(glmm_gauss2_log, ~ . - (1 | observer))
# compare with anova:
anova(glmm_gauss2_log, glmm_no_observer)
```

As with parks, including observer as a random effect improves model fit significantly ($p = 2.88e-6$), indicating meaningful body mass variation across observers within the model. 


## 12. Conclusion
Distance to water was a significant factor on giraffe body mass with a 1.2% increase in body mass for every km further from water a giraffe is observed in this model ($p = 0.025$), thus $H_0$ may be rejected at the 5% significance level. Increased body mass further from a water source may be a result of reduced foraging competition away from busy water sites, or of unrelated seasonal changes in body mass paired with different seasonal distributions of water sites. Without temporal data included here it is hard to disentangle.

Additional meaningful variations in body mass factors were observed in the park and observer random effects. Zebra Horizon Park masses were recorded ~10% higher than the expected population value; masses from observer 1 were ~10% higher, and observer 2 ~9% lower, than the expected population value. While these random effects are not statistically significant in terms of $p$ value, they indicate a contribution to baseline body mass variation in the population from different random effect levels. The variance in the response explained by these random effects also far outweighed those from the fixed effects (19.5% vs. 2.7%), which could indicates potential biases or differences in the methods used to measure body mass between observers, as well as demonstrating the limited power of the fixed effects in explaining body mass in this model.

As such a large amount of the model's variance was unexplained (77.9%), it means that the majority of the prediction power comes from unexplained variance, and the current predictors have a relatively weak predictive power. This could be due to missing other important predictors in the data such as: temporal data, temperatures, population counts, elevation data, type of habitat, and vegetation density and quality. If these predictor data were included, as well changes from binary to more detailed descriptions for `landscape` and `vegetation`, they could reduce residual variance, and effectively move these predictor effects out of the unexplained variance $\epsilon_i$, and into the fixed effects. Increasing the predictive abilities of the model, and in turn the precision of the model, would not necessarily uncover any other significant predictors but it could shift their coefficients, standard errors, and $p$ values due to new interactions. In turn this could help reveal significant predictors which are currently stored as "noise" within the unexplained variance, or even identify those currently in the fixed effects of the model as significant. 

Overall, distance to water has a small but significant effect on giraffe body mass, but most of the observed variation remains unexplained or associated with potential biases in random effects.








